### 2020-02-23 analysis/paper.Rmd
add methods to rmd

---
title: "A replication of a simulation study on measuring and correcting bias of under-5 mortality estimation in populations affected by HIV/AIDS"
author: "Sihao Miao, Ruian Yang, Xiaolu Qian, Jiyu Wang, David Wei"
date: "`r format(Sys.time(), '%d %B, %Y')`"
Repository:
  article:   
  code:      
  data:      
  notebook:  
reproduction:
  - "Quattrochi, John, et al. “Measuring and Correcting Bias in Indirect Estimates of under-5 Mortality in Populations Affected by HIV/AIDS: a Simulation Study.” BMC Public Health, vol. 19, no. 1, Dec. 2019, doi:10.1186/s12889-019-7780-3."
bibliography:
  bibliography.bib
output:
  output: bookdown::word_document2
---

# Introduction
Generalized linear model (GLM) is a widely used method in research. It is well-established with high flexibility and relatively straightforward interpretation. However, GLMs usually require making assumptions and having the underlying assumptions not hold could invalidate the whole model.

The application of GLM can be observed in many fields including public health. @quattrochi2019 introduced an GLM via penalized maximum likelihood to correct the under-5 mortality (U5M) in populations affected by HIV/AIDS. This method was developed since the U5M is commonly estimated using survey-based approaches in less developed regions and it is based on a fundamental assumption that the mother's survival is not related to her children's, which is often violated in populations affected by HIV/AIDS. Therefore, the author sought a method to correct this bias using GLM.

In our work, we present our endeavor to replicate the model building and regression analysis introduced in @quattrochi2019. In the meantime, we attempted using different parameters in both constructing the model and applying it to test how robust the original method was. We will in this paper report our method and results from the replication attempt.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Methods

```{r, load-packages}

```

## Overview
The original model was implemented in R (v3.5.2) [team2013r] with web application (https://johnquattrochi.com/bias/) to show that indirect method such as surveying can underestimate U5M by 0-41% in populations where 0-40% of the people are affected by HIV/AIDS. The model was written using the R language [team2013r] and some of the main packages used were tidyverse (v1.2.1) and foreach (v1.4.4). The original implementation utilized the doReids package (v1.1.1) that offers a back-end server for parallel processing that was not adopted in our replication due to the limitation in resources. In order to complete the simulation for constructing the model in a reasonable time frame without parallel computing, we reduced the number of populations used in the simulation input from 4480 to 1000. The code provided the author on Github (https://github.com/jquattro/hiv-childmort-bias) was mostly well-organized and documented. However, we discovered some mistakes regarding output inconsistency and wrong variable names. Despite the debugging process, we were able to replicate the implementation.

## Model Description

## Simulation Construction

## Model Fitting and Model Selection

Newly-generated simulation data was then imported and fitted to various models to select for a best-performing model which provides least errors.

The simulated data was first divided into training set (80%) and test set (20%). The training set was fitted to eight different models, which were further evaluated by calculating several error measures (root mean square error, root median square error, mean relative error and median relative error) on both training set (in-sample) and test set (out-of-sample) shown in Table 4.

The glmnet with alpha = 1 was selected as the best-performing model based on its error measures and was used in predictions later.

```{r, model-fitting-and-selection, include=FALSE}

# Get data from simulation
load("./results/regdata_all.Rdata")
load("./results/inputs.RData")

# add agegroup variable
nbd2k %<>% mutate(agegroup=rep(c(1:7),nrow(nbd2k)/7))

# Generate abs error and rel error
nbd2k %<>% mutate(abs.err=fiveq0_surv-fiveq0_hiv,
                  rel.err=(fiveq0_surv-fiveq0_hiv)/fiveq0_hiv
)

# new dep var 
nbd2k %<>% mutate(corr=fiveq0_hiv-fiveq0_surv,
                  corr2=corr+2.100105e-05)

##################################################################
##### MODEL FITTING AND MODEL SELECTION #####
##################################################################

compute_error_measures <- function(y,prediction,method,sample){

  # Absolute error
  abs.error <- abs(y-prediction)

  # Relative error
  relative.error <- abs.error/abs(y)

  # Root mean square error

  rmse <- sqrt(mean(abs.error^2))

  # Root median square error

  rmedse <- sqrt(median(abs.error^2))

  # Mean relative error

  mre <- mean(Filter(function(x) !is.infinite(x),relative.error),na.rm=TRUE)

  # Median relative error

  medre <- median(relative.error)

  data.frame(method=method,sample=sample,rmse=rmse,rmedse=rmedse,mre=mre,medre=medre,stringsAsFactors = FALSE)

}

# Prepare data
to.model <- nbd2k %>% mutate(
  agegroup=as.factor(agegroup))

full.model.formula <- corr ~ fiveq0_surv + agegroup + agegroup*hiv1990 + agegroup*hiv2000+  agegroup*hiv2010 + 
  agegroup*art_prev2005+ agegroup*art_prev2007+ agegroup*art_prev2009 + tfr2000 + tfr2010

minimum.model.formula <- corr ~ fiveq0_surv

# Out of sample data
fullsamp <- unique(to.model$i)
set.seed(400)

# dataset with 80% of full sample
ns1 <- sample(fullsamp, length(fullsamp)*.8, replace=FALSE)

# Select those i's from nbd2k
train.sample <- to.model %>% filter(i %in% ns1)

# dataset with 20% of full sample
tw1 <- setdiff(fullsamp,ns1)
test.sample <- to.model %>% filter(i %in% tw1)

error_table <- data.frame()

##### Full linear model #####

fit <- lm(full.model.formula,data=train.sample)
prediction <- predict(fit,newx = train.sample)
error_table %<>% bind_rows(compute_error_measures(train.sample$corr,prediction,"Full Linear Model","in-sample"))

prediction <- predict(fit,newx = test.sample)
error_table %<>% bind_rows(compute_error_measures(test.sample$corr,prediction,"Full Linear Model","out-of-sample"))

##### Forward and Backward selection #####

# maximum model

model.max <- glm(formula=full.model.formula,
                 family = "gaussian"(link="identity"), data=train.sample)

# minimum model

model.min <- glm(formula=minimum.model.formula, family = "gaussian"(link="identity"), data=train.sample)

# Forward BIC

fit <- step(object=model.min, scope=list(upper=model.max,lower=~1), direction="forward",k=log(nrow(train.sample)))

prediction <- predict(fit,newx = train.sample)

error_table %<>% bind_rows(compute_error_measures(train.sample$corr,prediction,"Forward Sel. BIC","in-sample"))

prediction <- predict(fit,newx = test.sample)

error_table %<>% bind_rows(compute_error_measures(test.sample$corr,prediction,"Forward Sel. BIC","out-of-sample"))

# Forward AIC

fit <- step(object=model.min, scope=list(upper=model.max,lower=~1), direction="forward",k=2)

prediction <- predict(fit,newx = train.sample)

error_table %<>% bind_rows(compute_error_measures(train.sample$corr,prediction,"Forward Sel. AIC","in-sample"))

prediction <- predict(fit,newx = test.sample)

error_table %<>% bind_rows(compute_error_measures(test.sample$corr,prediction,"Forward Sel. AIC","out-of-sample"))

# Backward BIC

fit <- step(model.max, direction="backward",k=log(nrow(to.model)))

prediction <- predict(fit,newx = train.sample)

error_table %<>% bind_rows(compute_error_measures(train.sample$corr,prediction,"Backward Sel. BIC","in-sample"))

prediction <- predict(fit,newx = test.sample)

error_table %<>% bind_rows(compute_error_measures(test.sample$corr,prediction,"Backward Sel. BIC","out-of-sample"))

# Backward AIC

fit <- step(model.max, direction="backward",k=2)

prediction <- predict(fit,newx = train.sample)

error_table %<>% bind_rows(compute_error_measures(train.sample$corr,prediction,"Backward Sel. AIC","in-sample"))

prediction <- predict(fit,newx = test.sample)

error_table %<>% bind_rows(compute_error_measures(test.sample$corr,prediction,"Backward Sel. AIC","out-of-sample"))

##### Glmnet #####
set.seed(500)
glmnet.errors <- lapply(c(0,0.5,1.0),function(alpha){

  # Get output vector for glmnet
  y <- train.sample %>% pull(all.vars(full.model.formula)[1])

  # Get predictors for glmnet
  x <- model.matrix(full.model.formula,train.sample)[,-1]

  set.seed(500)
  fit <- cv.glmnet(x,y,family="gaussian",alpha=alpha)

  prediction <- predict(fit,newx = x,s = "lambda.min")

  train.error <- compute_error_measures(train.sample$corr,prediction,paste0("glmnet, alpha=",alpha),"in-sample")

  xnew <- model.matrix(full.model.formula,test.sample)[,-1]

  prediction <- predict(fit,newx = xnew,s = "lambda.min")

  test.error <- compute_error_measures(test.sample$corr,prediction,paste0("glmnet, alpha=",alpha),"out-of-sample")

  bind_rows(train.error,test.error)

}) %>% bind_rows()

error_table %<>% bind_rows(glmnet.errors)

##### PCR #####

set.seed(100)
fit <- pcr(full.model.formula,data=train.sample,scale=TRUE,validation="CV", jackknife=TRUE)

ncomps <- c(20,30,selectNcomp(fit,method="onesigma",plot=TRUE)) %>% sort

pcr_errors <- lapply(ncomps,function(ncomp){

  prediction <- predict(fit,newdata = train.sample, type = "response",ncomp = ncomp)

  train.error <- compute_error_measures(train.sample$corr,prediction,paste0("PCR, ncomp=",ncomp),"in-sample")

  prediction <- predict(fit,newx = test.sample,ncomp = ncomp)

  test.error <- compute_error_measures(test.sample$corr,prediction,paste0("PCR, ncomp=",ncomp),"out-of-sample")

  bind_rows(train.error,test.error)

}) %>% bind_rows()

error_table %<>% bind_rows(pcr_errors)

##### PLS #####

set.seed(100)
fit <- plsr(full.model.formula,data=train.sample,scale=TRUE,validation="CV", jackknife=TRUE)
summary(fit)
validationplot(fit)

# 32 components are enough to account for at least 90% of variaiblity in both X and Y. Also use optimal number of components

ncomps <- c(32,selectNcomp(fit,method="onesigma",plot=TRUE)) %>% sort

pls_errors <- lapply(ncomps,function(ncomp){

  prediction <- predict(fit,newdata = train.sample, type = "response",ncomp = ncomp)

  train.error <- compute_error_measures(train.sample$corr,prediction,paste0("PLS, ncomp=",ncomp),"in-sample")

  prediction <- predict(fit,newx = test.sample,ncomp = ncomp)

  test.error <- compute_error_measures(test.sample$corr,prediction,paste0("PLS, ncomp=",ncomp),"out-of-sample")

  bind_rows(train.error,test.error)

}) %>% bind_rows()

error_table %<>% bind_rows(pls_errors)

###### Table 4 #####

# Prediction errors from models to correct bias in indirect estimates of U5M

# Save error table

ft <- error_table  %>% 
  mutate_at(vars(one_of("rmse","rmedse")),funs(sprintf("%0.6f",.))) %>%
  flextable() %>% set_header_labels(rmse="Root mean\nsquare error",
                                    rmedse="Root median\nsquare error",
                                    mre="Mean relative\nerror",
                                    medre="Median relative\nerror") %>%
  merge_v(j=1)

doc <- read_docx() %>%
  body_add_par(value = "Model selection.", style = "table title") %>% 
  body_add_flextable(ft)

doc %>% print("tables/table4.docx")

# glmnet with alpha = 1 was chosen to be the best-fitting-model

##### Best fitting model #####

# Get output vector for glmnet

y <- to.model %>% pull(all.vars(full.model.formula)[1])

# Get predictors for glmnet

x <- model.matrix(full.model.formula,to.model)[,-1]

alpha <- 1.0

set.seed(500)
best.fitting.model <- cv.glmnet(x,y,family="gaussian",alpha=alpha)

lambda <-best.fitting.model$lambda.min

# To use for predictions later.

best.fitting.model <- glmnet(x,y,family="gaussian",alpha=alpha,lambda = lambda)

save(x,best.fitting.model,full.model.formula,file="results/best_fitting_model.RData")

```

## Replication of Figures

```{r, ori-fig-4, fig.cap="Original figure 4"}
knitr::include_graphics("figures/figure4-ori.png")
```

```{r, ori-fig-4, fig.cap="Original figure 4"}
knitr::include_graphics("figures/figure5-ori.png")
```

The figure 4 (\@ref(fig:ori-fig-4)) and 5 (\@ref(fig:ori-fig-5)) in the original article were created by applying the best-performing model as well as an independent model (REF) to empirical data from 2010 DHS for Malawi and Tanzania respectively to estimate U5M of different age groups of population. By comparing differences in U5M estimates of unadjusted data and adjusted data by two models, the authors suggested that the new model was able to correct the bias from indirect estimates.

With the new simulation data and best-fitting model, we re-generated both figures by using R scripts provided (\@ref(fig:rep-fig-4) and \@ref(fig:rep-fig-5)).

To prepare data for plotting, we created three functions with R code provided by the authors to estimate U5M by indirect method, Ward-Zaba model and our best-performing model respectively. These three functions were then further integrated into a function called plot_fig to create our replicated figures.

```{r, indirect-estimates}

indirect_estimate <- function(dataset, fig_country){

  # extract data of the plotting country
  cntry <- dataset %>% filter(country==fig_country)

  # indirect estimates with empirical data
  source("R/indirect_estimates_functions_for_reg.R")
  ind_surv <- ind_est_computations(cntry,2010) %>% as.data.frame() %>% mutate(agegroup=1:7)
  ind_surv %<>% 
    select(agegroup,fiveq0=fiveq0, t_ref,refdate)

  # merge indirect estimates based on empirical data from Malawi or Tanzania with
  # data on HIV, ART and fertility from that same country
  # ind_est contains the results of indirect_estimates_functions applied to the data from Malawi or Tanzania
  # cntry contains the data from Malawi or Tanzania (from facevalidity.csv)
  iep <- cntry %>% 
    left_join(ind_surv,by="agegroup") %>% 
    rename(fiveq0_surv=fiveq0) %>% 
    mutate(agegroup=factor(agegroup)) %>% 
    select(-cd)
  colnames(iep) <- c("country","agegroup", "ceb","cs","tfr2010","tfr2000","hiv1990", "hiv2000", "hiv2010","art2005","art2006","art2007","art2008","art2009","art2010","art2011","art_prev2005","art_prev2006","art_prev2007","art_prev2008","art_prev2009","prop15to19_2010","prev15to19_2010","hiv2005","hiv2006","hiv2007","hiv2008","hiv2009","fiveq0_surv", "t_ref","refdate")

  return(iep) 
}

```

```{r, Ward-Zaba-model}
wz_model <- function(iep){
  # Ward & Zaba (2008) coefficients
  a=c(.1134,.1202,.1107,.1720,.1749,.1504,.2552)
  b=c(-.0226,-.0438,-.0882,-.1412,-.1758,-.1849,-.3269)
  c=c(-.1112,-.0978,-.0373,.0163,.042,.1807,.5594)

  # W&Z adjustment factor from applying coefficients to empirical data
  nz = a*iep$hiv2000+b*(iep$hiv2000^2)+c*iep$prev15to19_2010

  #Convert each n q 0 + n(z) into 5q0

  # Standard indirect estimates, adding Ward&Zaba adj factors
  # n = c(1,2,3,5,10,15,20)
  # UN General Female LT, e(0)=60, for above n
  l = c(.92688,.91025,.90151,.89193,.89534,.89002,.88195)
  q = 1-l
  logitMLT=.5*log(q/(1-q))
  alpha = NA
  fiveq0WZ = NA
  Y5 = NA

  nq0 <- ind_est_computations_core(cntry,2010)$nq0

  for(i in 1:7){
    Y5[i] = .5*log((nq0[i]+nz[i])/(1-(nq0[i]+nz[i])))
    alpha[i] = Y5[i] - logitMLT[i]
    fiveq0WZ[i] = exp(2*(alpha[i]+logitMLT[4]))/(1+exp(2*(alpha[i]+logitMLT[4])))
  }

  # fiveq0WZ are the Ward & Zaba estimates of U5M
  return(fiveq0WZ)
}
```

```{r, best-performing-model}
best_model <- function(iep, fig_name, best.fitting.model, full.model.formula){

  # Get the new predictions in matrix format
  nx <- model.matrix(full.model.formula,iep %>% mutate(corr=0))[,-1]

  # Make predictions
  prediction <- predict(best.fitting.model,newx = nx,se=TRUE,s="lambda.min") %>% 
    as.data.frame %>% 
    set_colnames("PredictedAdj")

  # Bootstrap confidence intervals for predictions
  if(is.null(boot_outs[[fig_name]])){
    # Get output vector for glmnet
    y <- to.model %>% pull(all.vars(full.model.formula)[1])

    # Get predictors for glmnet
    x <- model.matrix(full.model.formula,to.model)[,-1]
    Xy <-  cbind(y,x)
    boot_models <- list(length=5)
    t0 <- predict(best.fitting.model,newx = x,s="lambda.min") %>% as.vector
    tt <- boot_models[1:5] %>%purrr::map(~predict(best.fitting.model,newx = x,s="lambda.min") %>% as.vector) %>% unlist()
    #lambda <- best.fitting.model$lambda.min
    lambda <- best.fitting.model$lambda
    theta <- as.matrix(coef(best.fitting.model, s = "lambda.min"))
    pi_hat <- predict(best.fitting.model, newx = x, s = "lambda.min", type = "response")
    n <- length(pi_hat)
    y_star <- sapply(seq_len(length(tt)), function(i) ifelse(runif(n) <= pi_hat, 1, 0))
    suff_stat <- t(y_star) %*% x

    #bcapar(t0=t0,tt=tt,bb=suff_stat)
    # This is the function that we will use for each bootstrap sample
    theta <- function(xy,nx,glmnet_alpha,lambda){
      X <- xy[,2:(dim(xy)[2])]
      Y <- xy[,1]

      # Fit the model for the sample and using the same lambda as in the best model fitted
      m <- glmnet(X,Y,family="gaussian",alpha=glmnet_alpha,lambda=lambda)

      # Keep the models computed for each bootstrap sample in order to compute prediction intervals faster
      # boot_models <<- c(boot_models,list(m))

      # Prediction
      predict(m,newx = nx,s="lambda.min") %>% as.vector

    }

    # Keep the models computed for each bootstrap sample in order to compute prediction intervals faster
    # boot_models <- list()

    # Compute bootstrapped confedence intervals for each prediction point
    boot_outs[[fig_name]] <- lapply(1:nrow(nx),function(i){
      # i <- 1
      new_data <- nx[i,,drop=FALSE]

      set.seed(600)

      bcajack(x=Xy,B=500,func=theta,new_data,1.0,lambda)
    })
  }

  # Get two tailed 95% CI values 
  ci <- boot_outs[[fig_name]] %>% lapply(function(r){

    r$lims %>% as.data.frame() %>% mutate(quantile=as.numeric(rownames(.))) %>% filter(quantile %in% c(0.025,0.975)) %>% select(quantile,bca) %>% mutate(quantile=case_when(quantile==0.025 ~ "ci.lo",TRUE~ "ci.hi")) %>% spread(quantile,bca)
  }) %>% bind_rows()

  prediction %<>% bind_cols(ci)
  return(prediction)
}
```

```{r, figure-plot}
plot_fig <- function(fig_name, fig_country){
  # Load empirical data}
  fv <- fread("./data/facevalidity.csv") %>% set_colnames(c("country","agegroup", "ceb","cs","tfr2010","tfr2000","hiv1990", "hiv2000", "hiv2010","art2005","art2006","art2007","art2008","art2009","art2010","art2011","art_prev2005","art_prev2006","art_prev2007","art_prev2008","art_prev2009","prop15to19_2010","prev15to19_2010","hiv2005","hiv2006","hiv2007","hiv2008","hiv2009")) %>% mutate(cd=ceb-cs)

  # Load bootstrap estimates for model predictions
  if(file.exists("results/boot_outs.RData")){
    load("results/boot_outs.RData")
  }else{
    boot_outs <- list()  
  }

  # Load saved best fit model
  # Contains best.fitting.model and full.model.formula
  load("results/best_fitting_model.RData")

  # indirect estimates with empirical data
  iep <- indirect_estimate(fv, fig_country)

  # estimate from Ward & Zaba (2008) model
  fiveq0WZ <- wz_model(iep)

  # make prediction with best-performing model
  prediction <- best_model(iep, fig_name, best.fitting.model, full.model.formula)

  # Prepare data for plot
  to.plot <- bind_cols(iep, prediction,data.frame(fiveq0WZ=fiveq0WZ)) %>% 
    mutate(fit=fiveq0_surv+PredictedAdj,lwr=fiveq0_surv+ci.lo,upr=fiveq0_surv+ci.hi)%>% # Compute adjusted values and prediction interval upper and lower limits
    select(refdate,upr,lwr,Adjusted=fit,Unadjusted=fiveq0_surv,`Ward & Zaba`=fiveq0WZ) %>%   # Transform data for plot
    gather(type,value,-refdate,-upr,-lwr)

  # Plot
  ggplot(to.plot,aes(x=refdate,y=value,shape=type)) +
    geom_point(color="black", size=3) +
    geom_errorbar(aes(ymin=lwr, ymax=upr))+ 
    xlab("Year") +
    ylab("Under-five mortality (deaths per live birth)") +
    scale_shape_discrete(name="Estimate type", # Legend label, use darker colors
                         labels=c("Adjusted", "Unadjusted","Ward & Zaba"))+
    labs(title= fig_country) + 
    theme_classic() +
    theme(axis.text = element_text(color="black"))

}
```

```{r, rep-fig-4, fig.cap="Replication of figure 4 with new simulation data"}

# Figure 4
# Under-five mortality estimates using CEB/CS from 2010 DHS for Malawi:
# unadjusted (crude), adjusted (corrected) using best-performing model, and using Ward &
# Zaba (2008) model

plot_fig('figure4','Malawi')

```

```{r, rep-fig-5, fig.cap="Replication of figure 5 with new simulation data"}

# Figure 5
# Under-five mortality estimates using CEB/CS from 2010 DHS for Tanzania:
# unadjusted (crude), adjusted (corrected) using best-performing model, and using Ward &
# Zaba (2008) model

plot_fig('figure5','Tanzania')

```

# Result

# Conclusion

# References cited