### 2020-02-23 analysis/paper.Rmd
edit method and reference of the paper

---
title: "A replication of a simulation study on measuring and correcting bias of under-5 mortality estimation in populations affected by HIV/AIDS"
author: "Sihao Miao, Ruian Yang, Xiaolu Qian, Jiyu Wang, David Wei"
date: "`r format(Sys.time(), '%d %B, %Y')`"
Repository:
  article:   
  code:      
  data:      
  notebook:  
reproduction:
  - "Quattrochi, John, et al. “Measuring and Correcting Bias in Indirect Estimates of under-5 Mortality in Populations Affected by HIV/AIDS: a Simulation Study.” BMC Public Health, vol. 19, no. 1, Dec. 2019, doi:10.1186/s12889-019-7780-3."
bibliography:
  bibliography.bib
output:
  output: bookdown::word_document2
---

# Introduction
Generalized linear model (GLM) is a widely used method in research. It is well-established with high flexibility and relatively straightforward interpretation. However, GLMs usually require making assumptions and having the underlying assumptions not hold could invalidate the whole model.

The application of GLM can be observed in many fields including public health. @quattrochi2019 introduced an GLM via penalized maximum likelihood to correct the under-5 mortality (U5M) in populations affected by HIV/AIDS. This method was developed since the U5M is commonly estimated using survey-based approaches in less developed regions and it is based on a fundamental assumption that the mother's survival is not related to her children's, which is often violated in populations affected by HIV/AIDS. Therefore, the author sought a method to correct this bias using GLM.

In our work, we present our endeavor to replicate the model building and regression analysis introduced in @quattrochi2019. In the meantime, we attempted using different parameters in both constructing the model and applying it to test how robust the original method was. We will in this paper report our method and results from the replication attempt.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Methods

## Overview
The original model was implemented in R (v3.5.2) [team2013r] with web application (https://johnquattrochi.com/bias/) to show that indirect method such as surveying can underestimate U5M by 0-41% in populations where 0-40% of the people are affected by HIV/AIDS. The model was written using the R language [team2013r] and some of the main packages used were tidyverse (v1.2.1) and foreach (v1.4.4). The original implementation utilized the doReids package (v1.1.1) that offers a back-end server for parallel processing that was not adopted in our replication due to the limitation in resources. In order to complete the simulation for constructing the model in a reasonable time frame without parallel computing, we reduced the number of populations used in the simulation input from 4480 to 1000. The code provided the author on Github (https://github.com/jquattro/hiv-childmort-bias) was mostly well-organized and documented. However, we discovered some mistakes regarding output inconsistency and wrong variable names. Despite the debugging process, we were able to replicate the implementation.

## Model Description
The goal of this large-scale simulation used in the original paper is to build a wide range of demographic histories that accounts for the variability between actual populations, to measure how biases such as the independence assumption of mother and children survival rate in a population influenced by HIV/AIDS would exist. @quattrochi2019 used stochastic simulation methods to generate a set of population with input parameters such as demographic features, mortality rate for U5M, maternal mortality rate, HIV prevalence, etc. The input simulation parameters vary over a large interval for different age groups across different populations. In each time step of one year, each woman in the population has some probabilities to give birth for those who are 15 to 49 years old, being infected with HIV, initiating ART (if HIV-positive), and dying given the simulation parameters. Through the aging process in the simulation population, children born with infected HIV mothers will have a probability of getting infected while all children will have a probability of mortality.

## Simulation Construction 
### Range of Inputs

Since the primary goal in @quattrochi2019 was to measure bias in indirect estimates across a set of populations that have experienced different rates of fertility, mortality, HIV infection, and ART (antiretroviral therapy) initiation. We adopted the same method to generate a set of population using the following then inputs: fertility, adult mortality, U5M, percent of 15–19 year olds who are sexually active, maternal mortality in 1990, percent annual decline in the maternal mortality rate, HIV incidence, duration of breastfeeding, and ART coverage. The original paper simulated one population for each combination of inputs resulting in a total of 4480 populations. Since the original paper did not mention the amount of time required to complete the simulation, we started our process with the same 4480 populations but without parallel computing, we could not complete the simulation within the timeframe of our study and stopped the simulation when it reached 1000. For the purpose of replication, we set the seed to be a different value before running the simulation.

### Simulation Parameters

Initially, we attempted to change the year parameter and the population age parameter in the simulation construction process. We tried to change the year interval of our study from [1906, 2010] to [1906, 1966] as well as changing the population age interval from [0,120] to [0,75].  This turned out to be unsuccessful since the the above parameters were not only used in the simulation process but also in the following model selection and regression application. For instance, to estimate HIV prevalence, the model divides the population into 7 age groups, which cannot be achieved if the year or age parameter was reduced.
As a result, we used the same size and date for the initial population in the simulation. We initiated with 22,500 women who were aged 15 years in 1906, and ran the simulation through 2010.

For other parameters in simulation, we followed the original paper and used the same data sources. Annual probability of birth was defined as a function of calendar year and mother’s age and set to zero for women younger than 15 years and older than 49 years. Estimates of age-specific fertility rates (ASFR) from the United Nations Population Division’s World Fertility Data [@unitednations] was used and for years when ASFR were not available, the adjusted the nearest available ASFR using the interpolated estimates of the total fertility rate (TFR) from the United Nations Population Division’s World Population Prospects [@unitednations_2013] was used.

As mentioned in the original paper, according to [@chen2010fertility], women between 15–19 years old who were HIV-positive experienced higher ASFRs compared to HIV-negative women, depending if they were sexually active. Therefore the ASFR parameter was adjusted using the ratios estimated by [@chen2010fertility]. ASFR was also adjusted based on the studies that have found an increase in pregnancy following ART [@myer2010impact][@makumbi2011associations][@homsy2009reproductive]. The effect of ART on fertility likely depends on a variety of factors including age, cluster of differentiation 4 (CD4) count at initiation according to the original paper. For the simulation model, the assumption that the ASFR for 15–19 year olds is not affected by ART was made to further simplify the simulation.
Inputs relating to maternal mortality was calculated using MMR (maternal deaths per 100,000 births) in 1990 and the annual decline in MMR since 1990. The initial value of the MMR according to [@hogan2010maternal] was either 0.0012 or 0.012 and the annual rate of decline was set to 0 or 7.3% and calculated as:

For other inputs including the annual probability of HIV infection[@hogan2012spline][@heuveline2003hiv], the parameters governing CD4[@hallett2008estimating], ART coverage[@databank], the annual probability of death[@worldhealthorganization_2013] and the probability of mother-to-child transmission of HIV[@stover2008spectrum], we also used the same data sources and derivation methods.

## ...

## Result

## Conclusion

## References cited